---
title: "Initial Data Research"
author: "Niloofar Khalajzadeh"
date: "2023-10-29"
output: pdf_document
---


README PRAAT-output files

All files with .merged.txt are created with the Software PRAAT, they are simple tables created from so-called TexGrids.

Each files contains the data for one participant, performing one task, in either their L1 or L2.

The first column is the soundfile-ID:
AMGOPP1_L1_Task2.wav is the ID of the participant (AMGOPP1), the langueage (L1 or L2), and the task (Task1 or Task2).
Task1 in L1 is not the exact same task as Task1 in L2, but a very similar task requiring similar linguistic and cognitive demands.

The second column is "tmin": for duration-events, this is the start of the event; for point-events, the start and the end will be the same.

The third column is "tier": the name of the variable. These are the variables related to fluency:
Phrases: interval variable either 0 (speech) or 1 (silent pause)
DFauto (English/Dutch): interval variable which is 1 for a filled pause, and 0 for speech
Nuclei: point variable that is a number counting up, for each nucleus (middle) of a syllable throughout the sound file

The following tiers are all interval-tiers, related to the words, as looked up in an external corpus:
Lg10WF: Log frequency of occurrence in an external corpus (this is the best variable for lexical complexity)
SUBTLEXWF: Fequency per million in external corpus
CDcount: kind of freqency/count in external corpus
FREQlemma: yet another, frequency of the lemma not the word in external corpus
FREQlow: yet another frequency count
SUBTLEXCD: yet another frequency count
Lg10CD: log of a frequency count
CDlow: yet another frequency count
FREQcount: yet another frequency count
POStags: part of speech tag

The following two tiers are also interval-tiers, related to the words
FreqDist: measure for frequency of occurrence within the soundfile itself
Repetitions: measure for repetetiveness of the word until now. It ranges from 0 (word never uttered until now) to 1 (previous word was the same word)

The fourth column is the value of the tier (called 'text')
Phrases: 1 or 0
DFauto: 1 or 0
Nuclei: an absolute number
Lg10WF etc: a number
POStags: a code
FreqDist: a number
Repetitions: a number ranging from 0 to 1

NOTE that sometimes PRAAT outputs "--undefined--" or "MISSING", which you can change to "NA"

The fifth column is 'tmax', which is the same value of 'tmin' for the point tier "Nuclei", and which is the ending time of interval-tiers

# Loading packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(stringr)
library(readr)
library(dplyr)
library(patchwork)
library(ggplot2)
library(caret)
library(randomForest)
library(mgcv)
library(Metrics)
library(car)
```


```{r}
FILES_DIRECTORY <- "./files_converted/"
```


```{r}
get_soundtrack <- function() {
  recordings <- list()
  for (file in list.files(FILES_DIRECTORY)) {
    if (endsWith(file, ".merged.txt")) {
      recordings <- append(recordings, file)
    }
  }
  recordings
}
```


```{r message=FALSE, warning=FALSE, include=FALSE}
recordings <- get_soundtrack()

df <- data.frame()

for (r in recordings) {
  f_path <- paste(FILES_DIRECTORY, r, sep="")
  r_df <- read_table(f_path)
  df <- rbind(df, r_df)
}
df[df == "--undefined--"] <- NA
df[df == "MISSING"] <- NA

```


```{r}
SELECTED_PARTICIPANT_ID = "KIKIPP2"
TIME_INTERVAL = 5
```

## Prepare "dfauto_df" data frame
```{r include=FALSE}
dfauto_df <- df[startsWith(df$tier, "DFauto"),]
dfauto_df <- dfauto_df[!is.na(dfauto_df$text), ]

dfauto_df$tmin <- as.numeric(dfauto_df$tmin)
dfauto_df$tmax <- as.numeric(dfauto_df$tmax)

dfauto_df$duration <- dfauto_df$tmax - dfauto_df$tmin
dfauto_df$ParticipantID <- gsub("(.*)_L[12]_Task[12]\\.wav", "\\1", dfauto_df$SoundfileID)
dfauto_df$Language <- gsub(".*_(L[12])_Task[12]\\.wav", "\\1", dfauto_df$SoundfileID)
dfauto_df$Task <- gsub(".*_L[12]_(Task[12])\\.wav", "\\1", dfauto_df$SoundfileID)
dfauto_df$TransitionPoint <- dfauto_df$tmax
dfauto_df <- dfauto_df %>%
  group_by(ParticipantID, Language, Task) %>%
  mutate(PauseDuration = ifelse(text == 1, duration, 0)) %>%
  ungroup()

print(dfauto_df)
```

## Prepare "phrases_df" data frame
```{r include=FALSE}
phrases_df <- df[startsWith(df$tier, "Phrases"),]
phrases_df <- phrases_df[!is.na(phrases_df$text), ]

phrases_df$tmin <- as.numeric(phrases_df$tmin)
phrases_df$tmax <- as.numeric(phrases_df$tmax)

phrases_df$duration <- phrases_df$tmax - phrases_df$tmin
phrases_df$ParticipantID <- str_extract(phrases_df$SoundfileID, "^[A-Z]+\\d+")
phrases_df$Language <- ifelse(str_detect(phrases_df$SoundfileID, "_L1_"), "L1", "L2")
phrases_df$Task <- ifelse(str_detect(phrases_df$SoundfileID, "_Task1"), "Task1", "Task2")
phrases_df$TransitionPoint <- phrases_df$tmax
phrases_df <- phrases_df %>%
  group_by(ParticipantID, Language, Task) %>%
  mutate(PauseDuration = ifelse(text == 1, duration, 0)) %>%
  ungroup()

print(phrases_df)
```

## Prepare "nuclei_df" data frame
```{r include=FALSE}
nuclei_df <- df[startsWith(df$tier, "Nuclei"),]
nuclei_df <- nuclei_df[!is.na(nuclei_df$text), ]

nuclei_df$tmax <- as.double(nuclei_df$tmax)

nuclei_df$ParticipantID <- gsub("(.*)_L[12]_Task[12]\\.wav", "\\1", nuclei_df$SoundfileID)
nuclei_df$Language <- gsub(".*_(L[12])_Task[12]\\.wav", "\\1", nuclei_df$SoundfileID)
nuclei_df$Task <- gsub(".*_L[12]_(Task[12])\\.wav", "\\1", nuclei_df$SoundfileID)

print(nuclei_df)
```

## Prepare "wordfreq_df" data frame
```{r include=FALSE}
wordfreq_df <- df[startsWith(df$tier, "Lg10WF"),]
wordfreq_df <- wordfreq_df[!is.na(wordfreq_df$text), ]

wordfreq_df$tmin <- as.numeric(wordfreq_df$tmin)
wordfreq_df$tmax <- as.numeric(wordfreq_df$tmax)

wordfreq_df$ParticipantID <- str_extract(wordfreq_df$SoundfileID, "^[A-Z]+\\d+")
wordfreq_df$Language <- ifelse(str_detect(wordfreq_df$SoundfileID, "_L1_"), "L1", "L2")
wordfreq_df$Task <- ifelse(str_detect(wordfreq_df$SoundfileID, "_Task1"), "Task1", "Task2")

print(wordfreq_df)
```


```{r}
dfauto_L1_Task1 <- dfauto_df[dfauto_df$Language == "L1" & dfauto_df$Task == "Task1", ]

dfauto_L1_Task1$ypos <- seq_along(dfauto_L1_Task1$tmin)

plot_L1_Task1 <-  ggplot(dfauto_L1_Task1, aes(x = tmin, xend = tmax, y = SoundfileID, yend = SoundfileID)) +
  geom_segment(aes(color = SoundfileID), linewidth = 1) +
  theme_minimal() +
  labs(title = "Pauses in Soundtracks",
       x = "Time (seconds)",
       y = "Soundtrack ID")
```


```{r}
dfauto_L1_Task2 <- dfauto_df[dfauto_df$Language == "L1" & dfauto_df$Task == "Task2", ]

dfauto_L1_Task2$ypos <- seq_along(dfauto_L1_Task2$tmin)

plot_L1_Task2 <- ggplot(dfauto_L1_Task2, aes(x = tmin, xend = tmax, y = SoundfileID, yend = SoundfileID)) +
  geom_segment(aes(color = SoundfileID), size = 1) +
  theme_minimal() +
  labs(title = "Pauses in Soundtracks",
       x = "Time (seconds)",
       y = "Soundtrack ID")
```


```{r}
dfauto_L2_Task1 <- dfauto_df[dfauto_df$Language == "L2" & dfauto_df$Task == "Task1", ]

dfauto_L2_Task1$ypos <- seq_along(dfauto_L2_Task1$tmin)

plot_L2_Task1 <- ggplot(dfauto_L2_Task1, aes(x = tmin, xend = tmax, y = SoundfileID, yend = SoundfileID)) +
  geom_segment(aes(color = SoundfileID), size = 1) +
  theme_minimal() +
  labs(title = "Pauses in Soundtracks",
       x = "Time (seconds)",
       y = "Soundtrack ID")
```


```{r}
dfauto_L2_Task2 <- dfauto_df[dfauto_df$Language == "L2" & dfauto_df$Task == "Task2", ]

dfauto_L2_Task2$ypos <- seq_along(dfauto_L2_Task2$tmin)

plot_L2_Task2 <- ggplot(dfauto_L2_Task2, aes(x = tmin, xend = tmax, y = SoundfileID, yend = SoundfileID)) +
  geom_segment(aes(color = SoundfileID), size = 1) +
  theme_minimal() +
  labs(title = "Pauses in Soundtracks",
       x = "Time (seconds)",
       y = "Soundtrack ID")
```


```{r}
plot_L1_Task2 <- plot_L1_Task2 + theme(legend.position = "none")
plot_L2_Task2 <- plot_L2_Task2 + theme(legend.position = "none")

combined_plot <- plot_L1_Task2 + plot_L2_Task2 

combined_plot_layout <- combined_plot + plot_layout(ncol = 2, nrow = 1)

combined_plot_layout
```


```{r}
participant_plots <- lapply(unique(phrases_df$ParticipantID), function(participant) {
  participant_df <- phrases_df %>% filter(ParticipantID == participant)
  
  ggplot(participant_df, aes(x = tmin, xend = tmax, y = as.factor(text), yend = as.factor(text))) +
    geom_segment(aes(color = as.factor(text)), size = 1) +
    scale_color_manual(values = c("0" = "blue", "1" = "red")) +
    facet_grid(Language ~ Task) +
    labs(title = paste("Phrases for Participant", participant),
         x = "Time (seconds)",
         y = "Phrase (0 = Speech, 1 = Silent Pause)",
         color = "Phrase") +
    theme_minimal()
})

for (plot in participant_plots) {
  print(plot)
}

```


```{r}
participant_plots <- lapply(unique(dfauto_df$ParticipantID), function(participant) {
  participant_df <- dfauto_df %>% filter(ParticipantID == participant)
  
  ggplot(participant_df, aes(x = tmin, xend = tmax, y = as.factor(text), yend = as.factor(text))) +
    geom_segment(aes(color = as.factor(text)), size = 1) +
    scale_color_manual(values = c("0" = "green", "1" = "purple")) +
    facet_grid(Language ~ Task) +
    labs(title = paste("Dfauto for Participant", participant),
         x = "Time (seconds)",
         y = "0 = Speech, 1 = Filled Pause",
         color = "Dfauto") +
    theme_minimal()
})

for (plot in participant_plots) {
  print(plot)
}

```


```{r}
t_dfauto_df <- dfauto_df
t_phrases_df <- phrases_df

t_phrases_df$text <- as.numeric(phrases_df$text)
t_dfauto_df$text <- as.numeric(dfauto_df$text) + 2

all_participants <- unique(c(t_dfauto_df$ParticipantID))

participant_plots_2 <- lapply(all_participants, function(participant) {
  phrases_participant_df <- t_phrases_df %>% filter(ParticipantID == participant)
  dfauto_participant_df <- t_dfauto_df %>% filter(ParticipantID == participant)
  
  combined_df <- rbind(phrases_participant_df, dfauto_participant_df)

  ggplot(combined_df, aes(x = tmin, xend = tmax, y = as.factor(text), yend = as.factor(text))) +
    geom_segment(aes(color = as.factor(text)), size = 1) +
    scale_color_manual(values = c("0" = "blue", "1" = "red", "2" = "green", "3" = "purple")) +
    facet_grid(Language ~ Task) +
    scale_y_discrete(breaks = c("0", "1", "2", "3"), labels = c("Phrase - Speech", "Phrase - Silent Pause", "DFAuto - Speech", "DFAuto - Filled Pause")) +
    labs(title = paste("Phrases and Dfauto for Participant", participant),
         x = "Time (seconds)",
         y = "Phrase/Dfauto (0,1 = Phrases, 2,3 = Dfauto)",
         color = "Phrase/Dfauto") +
    theme_minimal()
})

for (plot in participant_plots_2) {
  print(plot)
}
```


## Utterance Fluency Measurements

- Speech rate = (Number of syllables/total time)
- Pruned speech rate Articulation rate = (Number of syllables – number of non-fluent syllables)/total time
- Articulation rate = (Number of syllables/speaking time)
- Pace = (Number of stressed syllables/total time)
- Mean length of utterance = (Total speaking time/number of utterances or Number of syllables/number of utterances)
- Number of silent pauses (per minute) = (Number of silent pauses/total time or speaking time)
- Mean duration of silent pauses = (Pausing time/number of silent pauses)
- Phonation time ratio = (Speaking time/Total time)
- Number of filled pauses (per minute) = (Number of filled pauses/total time or speaking time)
- Number of repetitions (per minute) = (Number of repetitions/total time or speaking time)
- Number of repairs (per minute) = (Number of repairs and restarts/total time or speaking time)


```{r}
# Speech rate (Number of syllables/ total time)

calculate_speech_rate <- function(data, language, task) {
  
  filtered_data <- data %>%
    filter(Language == language, Task == task)
  
  speech_rates <- c()
  time_intervals <- seq(0, max(filtered_data$tmax), by = TIME_INTERVAL)
  
  total_nucleis <- 0
  for (i in 1:(length(time_intervals) - 1)) {
    start_time <- time_intervals[i]
    end_time <- time_intervals[i + 1]
  
    timespan <- filtered_data[filtered_data$tmax >= start_time & filtered_data$tmax <= end_time,]
    timespan_ordered <- timespan[order(timespan$tmax),]

    nuclei_count <- as.numeric(timespan[nrow(timespan),]$text) - total_nucleis

    if (length(nuclei_count) < 1) {
      speech_rates <- c(speech_rates, 0)
      next
    }
    
    rate <- nuclei_count / TIME_INTERVAL
    total_nucleis = total_nucleis + nuclei_count

    speech_rates <- c(speech_rates, rate)
  }

  return(data.frame(TimeInterval = time_intervals[-length(time_intervals)], SpeechRate = speech_rates))
}

languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

speech_rate_data <- data.frame()

for (language in languages) {
  for (task in tasks) {
    nuclei_df_participant <- nuclei_df %>% filter(ParticipantID == SELECTED_PARTICIPANT_ID)

    result <- calculate_speech_rate(nuclei_df_participant, language, task)
    result$Language <- language
    result$Task <- task
    speech_rate_data <- bind_rows(speech_rate_data, result)
  }
}

ggplot(speech_rate_data, aes(x = TimeInterval, y = SpeechRate, group = interaction(Language, Task), color = interaction(Language, Task))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Speech Rate Over Time",
    x = "Time Interval (seconds)",
    y = "Speech Rate"
  ) +
  theme_minimal() +
  facet_grid(Language ~ Task)

```


```{r}
# Articulation Rate (Number of syllables/ speaking time)

calculate_articulation_rate <- function(nuclei_data, phrases_data, language, task) {
  
  nuclei_filtered <- nuclei_data %>% filter(Language == language, Task == task)
  
  phrases_filtered <- phrases_data %>% filter(Language == language, Task == task)
  
  silent_pause_intervals <- phrases_filtered %>%
    filter(text == 1) %>%
    select(tmin, tmax) %>%
    arrange(tmin)
  
  speaking_time <- 0
  for (i in 1:(nrow(silent_pause_intervals) - 1)) {
    speaking_time <- speaking_time + (silent_pause_intervals$tmin[i + 1] - silent_pause_intervals$tmax[i])
  }
  
  total_time <- max(nuclei_filtered$tmax)
  number_of_syllables <- nrow(nuclei_filtered)
  
  time_intervals <- seq(0, total_time, by = TIME_INTERVAL)
  articulation_rates <- numeric(length(time_intervals))

  for (i in 1:length(time_intervals)) {
    start_time <- time_intervals[i]
    end_time <- start_time + TIME_INTERVAL
    
    syllables_in_interval <- nuclei_filtered %>%
      filter(tmax >= start_time & tmax <= end_time)
    
    syllables_count <- nrow(syllables_in_interval)
    
    if (speaking_time > 0) {
      articulation_rates[i] <- syllables_count / TIME_INTERVAL
    } else {
      articulation_rates[i] <- 0
    }
  }
  
  result <- data.frame(
    Language = language,
    Task = task,
    TimeInterval = time_intervals,
    ArticulationRate = articulation_rates
  )
  
  return(result)
}

languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

articulation_rate_data <- data.frame()

for (language in languages) {
  for (task in tasks) {
    nuclei_df_participant <- nuclei_df %>% filter(ParticipantID == SELECTED_PARTICIPANT_ID)
    phrases_df_participant <- phrases_df %>% filter(ParticipantID == SELECTED_PARTICIPANT_ID)

    result <- calculate_articulation_rate(nuclei_df_participant, phrases_df_participant, language, task)
    articulation_rate_data <- bind_rows(articulation_rate_data, result)
  }
}

ggplot(articulation_rate_data, aes(x = TimeInterval, y = ArticulationRate, group = interaction(Language, Task), color = interaction(Language, Task))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Articulation Rate Over Time",
    x = "Time Interval (seconds)",
    y = "Articulation Rate"
  ) +
  theme_minimal() +
  facet_grid(Language ~ Task)

```


```{r}
# Number of Silent Pauses
phrases_df_participant <- phrases_df %>% filter(ParticipantID == SELECTED_PARTICIPANT_ID)

calculate_number_of_silent_pauses <- function(phrases_data, language, task) {
  
  phrases_filtered <- phrases_data %>% filter(Language == language, Task == task)
  
  number_of_silent_pauses <- phrases_filtered %>%
    filter(text == 1) %>%
    group_by(interval = cut(tmax, breaks = seq(0, 120, by = TIME_INTERVAL))) %>%
    summarise(Count = n()) %>%
    ungroup()
  
  number_of_silent_pauses$SilentPauses <- number_of_silent_pauses$Count
  
  number_of_silent_pauses$Language <- language
  number_of_silent_pauses$Task <- task
  
  return(number_of_silent_pauses)
}

languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

number_of_silent_pauses_data <- data.frame()

for (language in languages) {
  for (task in tasks) {
    result <- calculate_number_of_silent_pauses(phrases_df_participant, language, task)
    number_of_silent_pauses_data <- bind_rows(number_of_silent_pauses_data, result)
  }
}


ggplot(number_of_silent_pauses_data, aes(x = interval, y = SilentPauses, color = interaction(Language, Task), group = 1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Number of Silent Pauses",
    x = "Time Interval (seconds)",
    y = "Number of Silent Pauses"
  ) +
  theme_minimal() +
  facet_grid(Language ~ Task) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```


```{r}
# Number of filled puases

dfauto_df_participant <- dfauto_df %>% filter(ParticipantID == SELECTED_PARTICIPANT_ID)

calculate_number_of_filled_pauses <- function(dfauto_data, language, task) {
  
  dfauto_filtered <- dfauto_data %>%
    filter(Language == language, Task == task)
  
  number_of_filled_pauses <- dfauto_filtered %>%
    filter(text == 1) %>%
    group_by(interval = cut(tmax, breaks = seq(0, 120, by = TIME_INTERVAL))) %>%
    summarise(Count = n()) %>%
    ungroup()
  
  number_of_filled_pauses$FilledPauses <- number_of_filled_pauses$Count
  
  number_of_filled_pauses$Language <- language
  number_of_filled_pauses$Task <- task
  
  return(number_of_filled_pauses)
}

languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

number_of_filled_pauses_data <- data.frame()

for (language in languages) {
  for (task in tasks) {
    result <- calculate_number_of_filled_pauses(dfauto_df_participant, language, task)
    number_of_filled_pauses_data <- bind_rows(number_of_filled_pauses_data, result)
  }
}


ggplot(number_of_filled_pauses_data, aes(x = interval, y = FilledPauses, color = interaction(Language, Task), group = 1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Number of Filled Pauses",
    x = "Time Interval (seconds)",
    y = "Number of Filled Pauses"
  ) +
  theme_minimal() +
  facet_grid(Language ~ Task) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

# Predicting filled or silent pauses for one participant in each language and task
## Random Forest - Silent Pauses

```{r eval=FALSE, include=FALSE}

silentpauses_df <- number_of_silent_pauses_data %>%
  group_by(Language, Task) %>%
  arrange(interval) %>%
  mutate(
    lag1 = lag(SilentPauses, 1),
    lag2 = lag(SilentPauses, 2)
  ) %>%
  ungroup() 

silentpauses_df <- na.omit(silentpauses_df)

silentpauses_df$Language <- as.factor(silentpauses_df$Language)
silentpauses_df$Task <- as.factor(silentpauses_df$Task)

train_index <- floor(0.8 * nrow(silentpauses_df))

train_data <- silentpauses_df[1:train_index, ]
test_data <- silentpauses_df[(train_index + 1):nrow(silentpauses_df), ]

rf_model <- randomForest(SilentPauses ~ lag1 + lag2 + Language + Task, data = train_data)

predictions <- predict(rf_model, newdata = test_data)

mse <- mean((predictions - test_data$SilentPauses)^2)

mae <- mean(abs(predictions - test_data$SilentPauses))

print("===== MSE =====")
print(mse)
print("===== MAE =====")
print(mae)
```


```{r eval=FALSE, include=FALSE}
silentpauses_df$interval <- as.character(silentpauses_df$interval)
silentpauses_df$interval_n <- as.numeric(str_extract(silentpauses_df$interval, "(?<=,)(\\d+)(?=\\])"))

latest_data <- silentpauses_df %>%
  group_by(Language, Task) %>%
  summarize(latest_interval = max(interval_n), 
            latest_SilentPauses = last(SilentPauses),
            lag1 = nth(SilentPauses, n()-1),
            lag2 = nth(SilentPauses, n()-2)) %>%
  ungroup()

predict_data <- latest_data %>%
  select(-latest_SilentPauses) %>%
  mutate(interval = paste0("(", latest_interval, ",", latest_interval + TIME_INTERVAL, "]"))

predictions <- predict(rf_model, newdata = predict_data)
predict_data$Predicted_SilentPauses <- predictions
predicted_data <- predict_data %>%
  mutate(interval = paste0("(", latest_interval + TIME_INTERVAL, ",", latest_interval + 2*TIME_INTERVAL, "]"),
         SilentPauses = Predicted_SilentPauses,
         Count = Predicted_SilentPauses,
         Type = 'Predicted') %>%
  select(-latest_interval, -Predicted_SilentPauses, -lag1, -lag2)

number_of_silent_pauses_data$Type <- "Observed"
combined_data <- rbind(number_of_silent_pauses_data, predicted_data)

ggplot(combined_data, aes(x = interval, y = SilentPauses, color = interaction(Language, Task), group = 1)) +
  geom_line(aes(color = interaction(Language, Task))) +
  geom_point(data = subset(combined_data, Type == 'Observed'), aes(color = interaction(Language, Task))) +
  geom_point(data = subset(combined_data, Type == 'Predicted'), color = "red", size = 4, shape = 17) +
  labs(
    title = "Number of Silent Pauses",
    x = "Time Interval (seconds)",
    y = "Number of Silent Pauses"
  ) +
  theme_minimal() +
  facet_grid(Language ~ Task) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

## Random Forest - Filled Pauses

```{r eval=FALSE, include=FALSE}
library(randomForest)

filledpauses_df <- number_of_filled_pauses_data %>%
  group_by(Language, Task) %>%
  arrange(interval) %>%
  mutate(
    lag1 = lag(FilledPauses, 1),
    lag2 = lag(FilledPauses, 2)
  ) %>%
  ungroup() 

filledpauses_df <- na.omit(filledpauses_df)

filledpauses_df$Language <- as.factor(filledpauses_df$Language)
filledpauses_df$Task <- as.factor(filledpauses_df$Task)

train_index <- floor(0.8 * nrow(filledpauses_df))

train_data <- filledpauses_df[1:train_index, ]
test_data <- filledpauses_df[(train_index + 1):nrow(filledpauses_df), ]

rf_model <- randomForest(FilledPauses ~ lag1 + lag2 + Language + Task, data = train_data)

predictions <- predict(rf_model, newdata = test_data)

mse <- mean((predictions - test_data$FilledPauses)^2)

mae <- mean(abs(predictions - test_data$FilledPauses))

print("===== MSE =====")
print(mse)
print("===== MAE =====")
print(mae)
```

# Prediction of pause duration

### Prepare global "Pauses" dataframe

```{r}
dfauto_pauses <- dfauto_df[dfauto_df$text == 1,] %>%
  mutate(PauseType = "Filled")
phrases_pauses <- phrases_df[phrases_df$text == 1,] %>%
  mutate(PauseType = "Silent")

dfauto_phrases <- dfauto_df[dfauto_df$text == 0,] %>%
  mutate(PhraseType = "Filled")
phrases_phrases <- phrases_df[phrases_df$text == 0,] %>%
  mutate(PhraseType = "Silent")

pauses <- rbind(dfauto_pauses, phrases_pauses)
phrases <- rbind(dfauto_phrases, phrases_phrases)

pauses <- pauses[with(pauses, order(SoundfileID, tmin)),]
phrases <- phrases[with(phrases, order(SoundfileID, tmin)),]

# Remove any pause from beginning of the tracks

all_participants <- sort(unique(c(pauses$ParticipantID)))
# all_participants <- c(SELECTED_PARTICIPANT_ID)
 languages <- c("L1", "L2")
 tasks <- c("Task1", "Task2")
 
 for (participant in all_participants) {
   for (language in languages) {
     for (task in tasks) {
       track_phrases <- phrases %>% filter(ParticipantID == participant, Language == language, Task == task)
       track_phrases <- track_phrases[order(track_phrases$tmin),]
 
       pauses <- pauses %>% filter(!(ParticipantID == participant & Language == language & Task == task & tmin <= track_phrases[1,]$tmin))
     }
   }
 }

```

Include text about how ARIMA is not suitable in our case:

The applicability and performance of ARIMA depend significantly on the nature of your time series data, including its stationarity, seasonality, and the presence of trends. If your dataset is not a simple univariate time series, or if the pause durations are influenced by multiple factors (like the features you've created), you might need to explore multivariate time series models or stick with machine learning approaches that can handle multiple predictors.

```{r eval=FALSE, include=FALSE}
library(forecast)

pauses_durations <- pauses %>% filter(ParticipantID == SELECTED_PARTICIPANT_ID, Language == "L1", Task == "Task1")

pause_ts <- ts(pauses_durations$PauseDuration, frequency=1) # Adjust frequency as needed

fit <- auto.arima(pause_ts)
summary(fit)
forecasts <- forecast(fit, h=5)
print(forecasts)
plot(forecasts)
```

## Something like this for comparing Random Forest and GAM models in pause duration prediction

```{r eval=FALSE, include=FALSE}

all_participants <- sort(unique(c(pauses$ParticipantID)))
# all_participants <- c("AMGOPP1")
# all_participants <- c(SELECTED_PARTICIPANT_ID)

languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

for (participant in all_participants) {
  for (language in languages) {
    for (task in tasks) {
      track_pauses <- pauses %>% filter(ParticipantID == participant, Task == task, Language == language)
      track_pauses <- track_pauses[order(track_pauses$tmin),]

      track_phrases <- phrases %>% filter(ParticipantID == participant, Task == task, Language == language, PhraseType == "Filled")
      track_phrases <- track_phrases[order(track_phrases$tmin),]

      track_wordfreq <- wordfreq_df %>% filter(ParticipantID == participant, Task == task, Language == language)
      track_wordfreq <- track_wordfreq[order(track_wordfreq$tmin),]

      if (nrow(track_pauses) == 0 || nrow(track_phrases) == 0 || nrow(track_wordfreq) == 0) next
      
      pauses_5secs <- track_pauses %>%
        mutate(IntervalStart = floor(tmin / 5) * 5) %>%
        group_by(IntervalStart) %>%
        summarise(
          TotalDuration = sum(duration),
          Type = "Pauses"
        )

      
      X1_num_pauses_5_sec = c()
      X2_num_pauses_10_sec = c()
      X3_num_phrases_5_sec = c()
      X4_num_phrases_10_sec = c()
      X5_last_5_sec_pause_duration = c()
      X6_sum_word_freq_5_sec = c()
      
      for (row in 1:nrow(pauses_5secs)) {
        if (row == 1) {
          X1_num_pauses_5_sec = append(X1_num_pauses_5_sec, 0)
          X2_num_pauses_10_sec = append(X2_num_pauses_10_sec, 0)
          X3_num_phrases_5_sec = append(X3_num_phrases_5_sec, 0)
          X4_num_phrases_10_sec = append(X4_num_phrases_10_sec, 0)
          X5_last_5_sec_pause_duration = append(X5_last_5_sec_pause_duration, 0)
          X6_sum_word_freq_5_sec = append(X6_sum_word_freq_5_sec, 0)
          next
        }
        
        row_tmin <- pauses_5secs[row,]$IntervalStart

        X1_last_pause = append(X1_last_pause, track_pauses[row,]$tmin - track_pauses[row - 1,]$tmin)
        X2_last_pause_length = append(X2_last_pause_length, track_pauses[row - 1,]$duration)
        
        syllables_between <- track_nuclei %>% filter(tmin > track_pauses[row - 1,]$tmin, tmin < track_pauses[row,]$tmin)
        X3_syllables_since_last_pause = append(X3_syllables_since_last_pause, nrow(syllables_between))
        
        pauses_in_5_sec <- track_pauses %>% filter(tmin > row_tmin - 5, tmin < row_tmin)
        X1_num_pauses_5_sec = append(X1_num_pauses_5_sec, nrow(pauses_in_5_sec))

        pauses_in_10_sec <- track_pauses %>% filter(tmin > row_tmin - 10, tmin < row_tmin)
        X2_num_pauses_10_sec = append(X2_num_pauses_10_sec, nrow(pauses_in_10_sec))
        
        phrasess_in_5_sec <- track_phrases %>% filter(tmin > row_tmin - 5, tmin < row_tmin)
        X3_num_phrases_5_sec = append(X3_num_phrases_5_sec, nrow(phrasess_in_5_sec))
        
        phrasess_in_10_sec <- track_phrases %>% filter(tmin > row_tmin - 10, tmin < row_tmin)
        X4_num_phrases_10_sec = append(X4_num_phrases_10_sec, nrow(phrasess_in_10_sec))
        
        X5_last_5_sec_pause_duration = append(X5_last_5_sec_pause_duration, pauses_5secs[row - 1,]$TotalDuration)
        
        wordfreq_in_5_sec <- track_wordfreq %>% filter(tmin > row_tmin - 5, tmin < row_tmin)
        X6_sum_word_freq_5_sec = append(X6_sum_word_freq_5_sec, mean(as.numeric(wordfreq_in_5_sec$text)))
      }
      
      pauses_5secs$X1_num_pauses_5_sec <- X1_num_pauses_5_sec
      pauses_5secs$X2_num_pauses_10_sec <- X2_num_pauses_10_sec
      pauses_5secs$X3_num_phrases_5_sec <- X3_num_phrases_5_sec
      pauses_5secs$X4_num_phrases_10_sec <- X4_num_phrases_10_sec
      pauses_5secs$X5_last_5_sec_pause_duration <- X5_last_5_sec_pause_duration
      pauses_5secs$X6_sum_word_freq_5_sec <- X6_sum_word_freq_5_sec
      
      pauses_5secs <- na.omit(pauses_5secs)

      ###################################### GAM ######################################
      # Fit a GAM model
      gam_model <- gam(TotalDuration ~
                         X1_num_pauses_5_sec +
                         X2_num_pauses_10_sec +
                         X3_num_phrases_5_sec +
                         X4_num_phrases_10_sec +
                         X5_last_5_sec_pause_duration +
                         X6_sum_word_freq_5_sec
                         # s(X1_num_pauses_5_sec, k = 4, bs='ps', sp = 0.6) +
                         # s(X2_num_pauses_10_sec, k = 4, bs='ps', sp = 0.6) +
                         # s(X3_num_phrases_5_sec, k = 4, bs='ps', sp = 0.6),
                         # s(X4_num_phrases_10_sec, k = 4),
                         # s(X5_last_5_sec_pause_duration, sp=0.6, k = 4),
                      , data = pauses_5secs, family = gaussian(link = "log"))

      # Step 2: Make predictions on the existing dataset
      pauses_5secs$PredictedDuration <- predict(gam_model, newdata = pauses_5secs, type = "response")
      
      
      ###################################### GAM + CV ######################################
      gam_model_cv <- train(TotalDuration ~
                         X1_num_pauses_5_sec +
                         X2_num_pauses_10_sec +
                         X3_num_phrases_5_sec +
                         X4_num_phrases_10_sec +
                         X5_last_5_sec_pause_duration +
                         X6_sum_word_freq_5_sec, 
                  data = pauses_5secs,
                  method = "gam",
                  trControl = trainControl(method = "cv", number = 10),
                  tuneGrid = data.frame(method = "GCV.Cp", select = FALSE)
      )
      
      # Use the trained model to make predictions
      pauses_5secs$PredictedDurationGAM_CV <- predict(gam_model_cv, newdata = pauses_5secs)
      
      # Calculate evaluation metrics
      print(postResample(pred = pauses_5secs$PredictedDurationGAM_CV, obs = pauses_5secs$TotalDuration))
      
      
      ###################################### Random Forest ######################################
      set.seed(123) 
      train_indices <- sample(1:nrow(pauses_5secs), size = 0.8*nrow(pauses_5secs))
      train_data <- pauses_5secs[train_indices, ]
      test_data <- pauses_5secs[-train_indices, ]
      rf_model <- randomForest(
                    TotalDuration ~
                      X1_num_pauses_5_sec +
                      X2_num_pauses_10_sec +
                      X3_num_phrases_5_sec +
                      X4_num_phrases_10_sec +
                      X5_last_5_sec_pause_duration +
                      X6_sum_word_freq_5_sec
                  , data = train_data)
      
      pauses_5secs$PredictedDurationRF <- predict(rf_model, newdata = pauses_5secs, type = "response")
      
      ###################################### Random Forest + CV ######################################
      train_control <- trainControl(method = "cv", number = 10)
      formula <- TotalDuration ~
        X1_num_pauses_5_sec +
        X2_num_pauses_10_sec +
        X3_num_phrases_5_sec +
        X4_num_phrases_10_sec +
        X5_last_5_sec_pause_duration +
        X6_sum_word_freq_5_sec

      # Train the model with cross-validation
      rf_model_cv <- train(
        formula,
        data = pauses_5secs,
        method = "rf",
        trControl = train_control
      )

      # Use the trained model to make predictions
      pauses_5secs$PredictedDurationRF_CV <- predict(rf_model_cv, newdata = pauses_5secs)
      
      # Calculate evaluation metrics
      print(postResample(pred = pauses_5secs$PredictedDurationRF_CV, obs = pauses_5secs$TotalDuration))
      
      # Calculate MSE
      mse <- mean((pauses_5secs$PredictedDurationRF - pauses_5secs$TotalDuration)^2)
      print(paste(participant, language, task, "Test MSE:", mse))
      
      # Calculate RMSE
      rmse <- sqrt(mse)
      print(paste(participant, language, task, "Test RMSE:", rmse))
      
      # Calculate MAE
      mae <- mean(abs(pauses_5secs$PredictedDurationRF - pauses_5secs$TotalDuration))
      print(paste(participant, language, task, "Test MAE:", mae))

      # Step 3: Visualize Actual vs. Predicted Durations
      plt <- ggplot(pauses_5secs, aes(x = as.factor(IntervalStart))) +
        geom_bar(aes(y = TotalDuration, fill = "Pause Duration"), stat = "identity", position = "dodge", color = "steelblue") +
        geom_line(aes(y = PredictedDuration, group = 1, color = "Pred - GAM"), size = 1) +
        geom_line(aes(y = PredictedDurationGAM_CV, group = 1, color = "Pred - GAM + CV"), size = 1) +
        geom_line(aes(y = PredictedDurationRF, group = 1, color = "Pred - RF"), size = 1) +
        geom_line(aes(y = PredictedDurationRF_CV, group = 1, color = "Pred - RF + CV"), size = 1) +
        scale_fill_manual(values = c("Pause Duration" = "steelblue")) +
        scale_color_manual(values = c("Pred - GAM" = "black", "Pred - GAM + CV" = "purple", "Pred - RF" = "red", "Pred - RF + CV" = "blue")) +
        theme_minimal() +
        labs(x = "Interval Start (s)", y = "Duration (s)", title = paste("Actual vs. Predicted Durations for Each Interval - ", participant, task, language)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

      print(plt)
      
    }
  }
}
```


```{r}
all_participants <- sort(unique(c(pauses$ParticipantID)))
# all_participants <- c(SELECTED_PARTICIPANT_ID)
  
languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

X1_prev_pause_duration = c()
X2_prev_speech_duration = c()
X3_prev_600_msec_complexity = c ()
X4_prev_speech_syllables = c()
X5_prev_pause_type = c()
X6_prev_speech_rate = c()
X7_next_600_msec_complexity = c()
#X8_prev_speech_complexity = c()
#X9_prev_articulation_rate = c()

for (participant in all_participants) {
  for (language in languages) {
    for (task in tasks) {
      track_pauses <- pauses %>% filter(ParticipantID == participant, Task == task, Language == language)
      track_pauses <- track_pauses[order(track_pauses$tmin),]
      
      track_phrases <- phrases %>% filter(ParticipantID == participant, Task == task, Language == language, tier == "Phrases")
      track_phrases <- track_phrases[order(track_phrases$tmin),]

      track_nuclei <- nuclei_df %>% filter(ParticipantID == participant, Task == task, Language == language)
      
      track_wordfreq <- wordfreq_df %>% filter(ParticipantID == participant, Task == task, Language == language)
      track_wordfreq <- track_wordfreq[order(track_wordfreq$tmin),]

      if (nrow(track_pauses) == 0 || nrow(track_phrases) == 0 || nrow(track_wordfreq) == 0) next

      for (row in 1:nrow(track_pauses)) {
        if (row == 1) {
          X1_prev_pause_duration = append(X1_prev_pause_duration, 0)
          X2_prev_speech_duration = append(X2_prev_speech_duration, 0)
          X3_prev_600_msec_complexity = append(X3_prev_600_msec_complexity, 0)
          X4_prev_speech_syllables = append(X4_prev_speech_syllables, 0)
          X5_prev_pause_type = append(X5_prev_pause_type, "Silent")
          X6_prev_speech_rate = append(X6_prev_speech_rate, 0)
          X7_next_600_msec_complexity = append(X7_next_600_msec_complexity, 0)
#         X8_prev_speech_complexity = append(X8_prev_speech_complexity, 0)
#         X9_prev_articulation_rate = append(X9_prev_articulation_rate, 0)
          next
        }

        X1_prev_pause_duration = append(X1_prev_pause_duration, track_pauses[row - 1,]$duration)
        
        X2_prev_speech_duration = append(X2_prev_speech_duration, track_pauses[row,]$tmin - track_pauses[row - 1,]$tmin)
        
        X3_prev_600_msec_complexity = append(X3_prev_600_msec_complexity, ifelse(nrow(wordfreq_prev_600_msec) > 0, max(as.numeric(wordfreq_prev_600_msec$text)), 0))
        
        syllables_between <- track_nuclei %>% filter(tmin > track_pauses[row - 1,]$tmin, tmin < track_pauses[row,]$tmin)
        X4_prev_speech_syllables = append(X4_prev_speech_syllables, nrow(syllables_between))

        X5_prev_pause_type = append(X5_prev_pause_type, track_pauses[row - 1,]$PauseType)
        
        syllables_so_far = track_nuclei %>% filter(tmin > 0, tmin < track_pauses[row,]$tmin)
        X6_prev_speech_rate = append(X6_prev_speech_rate, nrow(syllables_so_far) / track_pauses[row,]$tmin)
        
        speech_time_so_far = track_phrases %>% filter(tmin > 0, tmin < track_pauses[row,]$tmin) %>% filter(tier == "Phrases")
        
        wordfreq_next_600_msec = track_wordfreq %>% filter(tmin > track_pauses[row,]$tmax, tmin < track_pauses[row,]$tmax + 0.6)
        X7_next_600_msec_complexity = append(X7_next_600_msec_complexity, ifelse(nrow(wordfreq_next_600_msec) > 0, max(as.numeric(wordfreq_next_600_msec$text)), 0))

        wordfreq_prev_600_msec = track_wordfreq %>% filter(tmin < track_pauses[row,]$tmax, tmin > track_pauses[row,]$tmax - 0.6)
        
#        wordfreq_between <- track_wordfreq %>% filter(tmin > track_pauses[row - 1,]$tmin, tmin < track_pauses[row,]$tmin)
#        X8_prev_speech_complexity = append(X8_prev_speech_complexity, mean(as.numeric(wordfreq_between$text)))
        
#        X9_prev_articulation_rate = append(X9_prev_articulation_rate, nrow(syllables_so_far) / sum(speech_time_so_far$duration))
      }
    }
  }
}

pauses$X1_prev_pause_duration <- as.numeric(X1_prev_pause_duration)
pauses$X2_prev_speech_duration <- as.numeric(X2_prev_speech_duration)
pauses$X3_prev_600_msec_complexity <- as.numeric(X3_prev_600_msec_complexity)
pauses$X4_prev_speech_syllables <- as.numeric(X4_prev_speech_syllables)
pauses$X5_prev_pause_type <- as.factor(X5_prev_pause_type)
pauses$X6_prev_speech_rate <- as.numeric(X6_prev_speech_rate)
pauses$X7_next_600_msec_complexity <- as.numeric(X7_next_600_msec_complexity)
#pauses$X8_prev_speech_complexity <- as.numeric(X8_prev_speech_complexity %>% replace(is.na(.), 0))
#pauses$X9_prev_articulation_rate <- as.numeric(X_prev_articulation_rate)

pauses <- pauses[with(pauses, order(SoundfileID, tmin)),]
pauses <- na.omit(pauses)
```

# Random Forest

```{r eval=FALSE, include=FALSE}

all_participants <- c(SELECTED_PARTICIPANT_ID)
languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

for (participant in all_participants) {
  for (language in languages) {
    for (task in tasks) {
      track_pauses <- pauses %>%
                      filter(ParticipantID == participant, Task == task, Language == language) %>%
                      arrange(tmin)

      set.seed(123) 
      trainIndex <- createDataPartition(track_pauses$PauseDuration, p = .8, 
                                        list = FALSE, 
                                        times = 1)
      trainData <- track_pauses[trainIndex, ]
      testData <- track_pauses[-trainIndex, ]

      rf_model <- randomForest(PauseDuration ~ X1_prev_pause_duration + 
                                                X2_prev_speech_duration + 
                                                X4_prev_speech_syllables +
                                                X5_prev_pause_type +
                                                X6_prev_speech_rate +
                                                TransitionPoint +
                                                PauseType,
                                data = trainData, 
                                ntree = 500) 

      print(rf_model)
      
      predictions <- predict(rf_model, newdata = testData)
      testData$Predictions <- predictions
      
      
      importance_df <- as.data.frame(importance(rf_model))
      print(importance_df)
      importance_df$Feature <- rownames(importance_df)
      plt <- ggplot(importance_df, aes(x = reorder(Feature, IncNodePurity), y = IncNodePurity)) +
        geom_bar(stat = "identity") +
        coord_flip() +
        xlab("Feature") + ylab("Importance") +
        ggtitle(paste("Feature Importance for Participant", participant, "Language", language, "Task", task))
      print(plt)
      
      plt <- ggplot(testData, aes(x = tmin, y = Predictions)) +
        geom_point(alpha = 0.5) +
        geom_line(aes(x = tmin, y = PauseDuration), color = "red") +  
        labs(x = "Actual Pause Duration", y = "Predicted Pause Duration", 
             title = paste("Actual vs Predicted Pause Durations for Participant", participant, "Language", language, "Task", task))
      print(plt)
      
       rmse_value <- rmse(testData$PauseDuration, predictions)
       r_squared <- R2(predictions, testData$PauseDuration)
       cat("RMSE for Participant", participant, "Language", language, "Task", task, ":", rmse_value, "\n")
       cat("R^2 for Participant", participant, "Language", language, "Task", task, ":", r_squared, "\n")
    }
  }
}
```

# GAM Model

## 2 different approaches are taken in to account

### 1. One general model for each participant
# Adding new predictor previous and next 0.6 sec word complexity

```{r}

# all_participants <- sort(unique(c(pauses$ParticipantID)))
all_participants <- c(SELECTED_PARTICIPANT_ID)
  
languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

for (participant in all_participants) {
  participant_pauses <- pauses %>% filter(ParticipantID == participant)
  participant_pauses <- participant_pauses[order(participant_pauses$SoundfileID, participant_pauses$tmin),]
  
  set.seed(123) 
  trainIndex <- createDataPartition(participant_pauses$PauseDuration, p = .8,
                                    list = FALSE,
                                    times = 1)
  trainData <- participant_pauses[trainIndex, ]
  testData <- participant_pauses[-trainIndex, ]
  
      gam_model <- gam(PauseDuration ~ #s(X1_prev_pause_duration) + 
#                                         s(log1p(X2_prev_speech_duration)) + 
#                                         s(X3_prev_600_msec_complexity) +
#                                         s(X4_prev_speech_syllables) +
                                         factor(X5_prev_pause_type) +
                                         s(log1p(X6_prev_speech_rate)) +
                                         s(X7_next_600_msec_complexity) +
                                         s(TransitionPoint) +
                                         PauseType +
                                         log1p(X6_prev_speech_rate):PauseType +
#                                         Language:Task +
#                                         X1_prev_pause_duration:log1p(X6_prev_speech_rate) +
                                         X7_next_600_msec_complexity:Language:Task 
#                                         X7_next_600_msec_complexity:TransitionPoint 

                                         
                       ,data = trainData)

      print(summary(gam_model))

#      par(mfrow = c(3, 2))
      plot(gam_model, residuals = TRUE)

      predictions <- predict(gam_model, newdata = testData)
      testData$Predictions <- predictions
      testData$error <- testData$Predictions - testData$PauseDuration
      testData$errorType <- ifelse(testData$error >= 0, "Positive", "Negative")

      
      plt <- ggplot(testData, aes(x = tmin, y = Predictions)) +
        geom_point(alpha = 1, color = "red") +
        geom_point(aes(x = tmin, y = PauseDuration, color = PauseType)) +
        geom_segment(aes(xend = tmin, y = Predictions, yend = PauseDuration, color = errorType), size = 2, alpha = 0.5) +  
        labs(x = "time (s)", y = "Pause Duration (Actual vs Predicted)", title = "Actual vs Predicted Pause Durations") +
        theme_gray() +
        scale_color_manual(values = c("Filled" = "purple", "Silent" = "blue", "Positive" = "green", "Negative" = "red"))
      
      print(plt)

      rmse_value <- rmse(testData$PauseDuration, predictions)
      r_squared <- R2(predictions, testData$PauseDuration)
      
      cat("RMSE for Participant", participant, ":", rmse_value, "\n")
      cat("R^2 for Participant", participant, ":", r_squared, "\n")
      print("======================================================================")
    }

```

### TransitionPoint

```{r}
data_plot <- ggplot(data = trainData, aes(y = PauseDuration, x = TransitionPoint, colour = PauseType)) +
    geom_point() + geom_line(aes(y = fitted(gam_model)), colour = "lightgreen",
    size = 1, alpha = 0.5) + theme_bw()

data_plot
```

### Previous Speech Rate on log scale

```{r}
data_plot <- ggplot(data = trainData, aes(y = PauseDuration, x = log1p(X6_prev_speech_rate), colour = PauseType)) +
    geom_point() + geom_line(aes(y = fitted(gam_model)), colour = "lightgreen",
    size = 1, alpha = 0.5) + theme_bw()

data_plot
```

### Nexr 0.6 sec word complexity

```{r}
data_plot <- ggplot(data = trainData, aes(y = PauseDuration, x = X7_next_600_msec_complexity, colour = PauseType)) +
    geom_point() + geom_line(aes(y = fitted(gam_model)), colour = "lightgreen",
    size = 1, alpha = 0.5) + theme_bw()

data_plot
```

```{r}
gam.check(gam_model)
```

```{r}
concurvity(gam_model, full = TRUE)
```

```{r}
concurvity(gam_model, full = FALSE)
```

### New Participant

```{r}
# all_participants <- sort(unique(c(pauses$ParticipantID)))
all_participants <- c("KIKIPP2")
  
languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

for (participant in all_participants) {
  participant_pauses <- pauses %>% filter(ParticipantID == participant)
  participant_pauses <- participant_pauses[order(participant_pauses$SoundfileID, participant_pauses$tmin),]
  
  set.seed(123) 
  trainIndex <- createDataPartition(participant_pauses$PauseDuration, p = .8,
                                    list = FALSE,
                                    times = 1)
  trainData <- participant_pauses[trainIndex, ]
  testData <- participant_pauses[-trainIndex, ]
  
      gam_model <- gam(PauseDuration ~ s(X1_prev_pause_duration) + 
#                                         s(log1p(X2_prev_speech_duration)) + 
#                                         s(X3_prev_600_msec_complexity) +
#                                         s(X4_prev_speech_syllables) +
                                         factor(X5_prev_pause_type) +
                                         s(log1p(X6_prev_speech_rate)) +
                                         s(X7_next_600_msec_complexity) +
                                         s(TransitionPoint) +
                                         PauseType +
                                         log1p(X6_prev_speech_rate):PauseType +
                                         Language:Task +
                                         X1_prev_pause_duration:log1p(X6_prev_speech_rate) 
#                                         X7_next_600_msec_complexity:Language 
#                                         X7_next_600_msec_complexity:TransitionPoint 

                                         
                       ,data = trainData)

      print(summary(gam_model))

#      par(mfrow = c(3, 2))
      plot(gam_model, residuals = TRUE)

      predictions <- predict(gam_model, newdata = testData)
      testData$Predictions <- predictions
      testData$error <- testData$Predictions - testData$PauseDuration
      testData$errorType <- ifelse(testData$error >= 0, "Positive", "Negative")

      
      plt <- ggplot(testData, aes(x = tmin, y = Predictions)) +
        geom_point(alpha = 1, color = "red") +
        geom_point(aes(x = tmin, y = PauseDuration, color = PauseType)) +
        geom_segment(aes(xend = tmin, y = Predictions, yend = PauseDuration, color = errorType), size = 2, alpha = 0.5) +  
        labs(x = "time (s)", y = "Pause Duration (Actual vs Predicted)", title = "Actual vs Predicted Pause Durations") +
        theme_gray() +
        scale_color_manual(values = c("Filled" = "purple", "Silent" = "blue", "Positive" = "green", "Negative" = "red"))
      
      print(plt)

      rmse_value <- rmse(testData$PauseDuration, predictions)
      r_squared <- R2(predictions, testData$PauseDuration)
      
      cat("RMSE for Participant", participant, ":", rmse_value, "\n")
      cat("R^2 for Participant", participant, ":", r_squared, "\n")
      print("======================================================================")
    }

```

### 2. 4 different GAM models for one participant in each language and task

```{r}

# all_participants <- sort(unique(c(pauses$ParticipantID)))
all_participants <- c(SELECTED_PARTICIPANT_ID)
  
languages <- c("L1", "L2")
tasks <- c("Task1", "Task2")

for (participant in all_participants) {
  participant_plts = c()

  for (language in languages) {
    for (task in tasks) {
      track_pauses <- pauses %>% filter(ParticipantID == participant, Task == task, Language == language)
      track_pauses <- track_pauses[order(track_pauses$tmin),]
      
      set.seed(123) 
      trainIndex <- createDataPartition(participant_pauses$PauseDuration, p = .8, 
                                        list = FALSE, 
                                        times = 1)
      trainData <- track_pauses[trainIndex, ]
      testData <- track_pauses[-trainIndex, ]

      gam_model <- gam(PauseDuration ~ #s(X1_prev_pause_duration, k = 9) + 
                                         #s(X2_prev_speech_duration, k = 3) + 
#                                         s(X3_prev_600_msec_complexity, k = 3) +
#                                         s(X4_prev_speech_syllables, k = 3) +
                                         factor(X5_prev_pause_type) +
                                         s(log1p(X6_prev_speech_rate), k = 3) +
                                         s(X7_next_600_msec_complexity, k = 3) +
                                         s(TransitionPoint) +
                                         PauseType +
                                         X6_prev_speech_rate:X7_next_600_msec_complexity 
                       ,data = trainData)

      print(summary(gam_model))

#      par(mfrow = c(3, 2))
      plot(gam_model, residuals = TRUE)

      predictions <- predict(gam_model, newdata = testData)
      testData$Predictions <- predictions
      testData$error <- testData$Predictions - testData$PauseDuration
      testData$errorType <- ifelse(testData$error >= 0, "Positive", "Negative")

      
      plt <- ggplot(testData, aes(x = tmin, y = Predictions)) +
        geom_point(alpha = 1, color = "red") +
        geom_point(aes(x = tmin, y = PauseDuration, color = PauseType)) +
        geom_segment(aes(xend = tmin, y = Predictions, yend = PauseDuration, color = errorType), size = 2, alpha = 0.5) +  
        labs(x = "time (s)", y = "Pause Duration (Actual vs Predicted)", title = "Actual vs Predicted Pause Durations") +
        theme_gray() +
        scale_color_manual(values = c("Filled" = "purple", "Silent" = "blue", "Positive" = "green", "Negative" = "red"))
      
      print(plt)

      rmse_value <- rmse(testData$PauseDuration, predictions)
      r_squared <- R2(predictions, testData$PauseDuration)
      
      cat("RMSE for Participant", participant, "in Language", language, "for Task", task, ":", rmse_value, "\n")
      cat("R^2 for Participant", participant, "in Language", language, "for Task", task, ":", r_squared, "\n")
      print("======================================================================")
    }
  }
}
```

# GAMM Model

```{r eval=FALSE, include=FALSE}
# Sample GAMM model fitting with crossed random effects
gamm_model <- bam(PauseDuration ~ s(X1_prev_pause_duration) + 
                                    s(X2_prev_speech_duration) + 
                                    s(X3_prev_speech_complexity) + 
                                    s(X4_prev_speech_syllables) +
                                    factor(X5_prev_pause_type) +
                                    s(tmin) +
                                    PauseType +
                                    Task
                  ,random = ~(1|ParticipantID) + (1|Language),data = pauses,
                  discrete = TRUE) # 'discrete = TRUE' can speed up large models

summary(gamm_model)

predicted_values <- predict(gamm_model, newdata = pauses)
pauses$PredictedPauseDuration <- predicted_values
actual_values <- pauses$PauseDuration


specific_soundtrack <- "AMGOPP1_L2_Task2.wav"
subset_data <- pauses[pauses$SoundfileID == specific_soundtrack, ]

subset_data_long <- reshape2::melt(subset_data, id.vars = "tmin", measure.vars = c("PauseDuration", "PredictedPauseDuration"))

ggplot(subset_data_long, aes(x = tmin, y = value, color = variable)) + 
  geom_line() + 
  labs(x = "Time", y = "Pause Duration", color = "Legend") +
  ggtitle(paste("Actual vs. Predicted Pause Durations for", specific_soundtrack)) +
  scale_color_manual(values = c("PauseDuration" = "blue", "PredictedPauseDuration" = "red"), labels = c("Actual", "Predicted")) +
  theme_minimal()

```